---
title: "README"
author: "J. Hartsfield"
date: "October 23, 2015"
output: html_document
---

## Get the data into R and load needed library dplyr
Go to the website, "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip" and download the file, then unzip the files. Read the files into R using read.table.

```{r getdata}
library(dplyr)
# read in the data
# create data directory if it does not exist
if(!file.exists("./data")){dir.create("./data")}
fileUrl <- "http://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(fileUrl, destfile="./data/temp.zip") # download file
unzip("./data/temp.zip", exdir = "./data") # unzip file
subject_test <- read.table("./data/UCI HAR Dataset/test/subject_test.txt")
X_test <- read.table("./data/UCI HAR Dataset/test/X_test.txt")
Y_test <- read.table("./data/UCI HAR Dataset/test/Y_test.txt")
subject_train <- read.table("./data/UCI HAR Dataset/train/subject_train.txt")
X_train <- read.table("./data/UCI HAR Dataset/train/X_train.txt")
Y_train <- read.table("./data/UCI HAR Dataset/train/Y_train.txt")
features <- read.table("./data/UCI HAR Dataset/features.txt")
activities <- read.table("./data/UCI HAR Dataset/activity_labels.txt")
```

## Combine the text files into one large file.
Using rbind and cbind, make one large array with subject, activity, and measurements of test observations followed by training observations.  Set appropriate column names from features vector. 
```{r combine}
Col1 <- rbind(subject_test,subject_train)
Col2 <- rbind(Y_test,Y_train)
dm <- cbind(Col1, Col2)
Col3 <- rbind(X_test, X_train)
dm <- cbind(dm, Col3)

# make a vector out of feature names and set as column names
M <- as.matrix(features)
M2 <- M[,2]
CN <- c("Subject","ActivityLabel",M2)
colnames(dm) <- CN
```
## Keep subset of columns that measure mean and std.
Keep only columns for subject number, activity label, and those for mean and std. Find column headings that have "mean" in them and exclude those with "meanFreq".  Keep those columns and those with "std".  Put this subset of the orginal data in an array called dm.  Replace the activity number with the name of the activity.
```{r subsetData}
keepind <- grepl("mean()", CN)
keepind <- keepind & ! grepl("meanFreq()",CN) # exclude meanFreq
keepind2 <- grepl("std()", CN)
keepind <- keepind | keepind2
keepind[1:2] <- TRUE # keep columns for subject and activity label
dm <- dm[,keepind]
# order the data by Subject and ActivityLabel
# dmord <- arrange(dm,Subject,ActivityLabel)

# replace ActivityLabel numbers with corresponding activity
for(i in 1:length(activities$V1)){
      dm$ActivityLabel <- gsub(activities$V1[i],activities$V2[i],dm$ActivityLabel)
}
```
## Output an array with the average of each measurement for each activity for each subject
Group by subject and activity label, then summarise each using the mean.  Write out the result using write.table.
```{r output}
df <- as.data.frame(dm)
df <- df %>% group_by(Subject,ActivityLabel) %>% summarise_each(funs(mean))
dm2 <- as.matrix(df)
write.table(dm2, file="run_analysis_output.txt", row.names = FALSE)
```